Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-18 arxiv/proof_examples.tex
section D wip

\paragraph{Part $1$: characterizing the global minimum of $f(A, B)$.}
Let $\hat{x}$ denote the global minimizer of $g(x)$.
We show that in the setting of Corollary \ref{cor_MTL_loss}, $\hat x$ is close to 1.
This gives us the global minimum of $f(A, B)$, since $\hat{B}$ is given by $\hat{x}$ using local optimality conditions.
First, we show that $g(x)$ and $\ex{g(x) \mid X^{(1)}, X^{(2)}}$ are close using standard concentration bounds.
\begin{claim}\label{claim_largedev1}
 In the setting of Corollary \ref{cor_MTL_loss}, for any $x = O(1)$, we have that with high probability
$$\left| g(x) - \ex{g(x) \mid X^{(1)}, X^{(2)} }\right| \le p^{1/2+c} \left(\sigma^2 + \kappa^2+d^2 \right).$$
\end{claim}
\begin{proof}
 There are two terms in $g(A)$ from equation \eqref{eq_mtl_A12}.
 We will focus on dealing with the concentration error of the first term. The second term is similar to the first and we omit the details.
 For the first term, we expand into several equations under various situations involving the random noise and the random-effect model.
 \begin{align}
 & \left\| X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (x\beta^{(2)}-\beta^{(1)})+ \left(x^2 X^{(1)}\hat \Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right)\epsilon^{(1)} \right. \nonumber\\
		& \left. + x X^{(1)}\hat \Sigma(x)^{-1} (X^{(2)})^\top \epsilon^{(2)} \right\|^2=h_1(x) + h_2(x) +h_3(x) + 2h_4(x) + 2h_5(x) + 2h_6(x), \label{expand_6}
 \end{align}
 where
 \begin{align*}
&h_1(x):=  (\beta^{(1)}-x\beta^{(2)})^\top (X^{(2)})^\top X^{(2)}  \hat\Sigma(x)^{-1}  (X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta^{(1)}-x\beta^{(2)})  ,\\
& h_2(x) := {(\epsilon^{(1)})^\top \left(x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right)^2\epsilon^{(1)}}  ,\\
& h_3(x):=  x^2 { (\epsilon^{(2)} )^\top X^{(2)} \hat\Sigma(x)^{-1}  (X^{(1)})^\top X^{(1)}\hat\Sigma(x)^{-1} (X^{(2)})^\top \epsilon^{(2)}} , \\
& h_4(x):=  (\epsilon^{(1)})^\top\left(x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right) X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (x\beta^{(2)}-\beta^{(1)}), \\
&h_5(x):=x (\epsilon^{(2)})^\top X^{(2)}\hat\Sigma(x)^{-1} (X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (x\beta^{(2)}-\beta^{(1)}),\\
&h_6(x):=x (\epsilon^{(2)})^\top X^{(2)}\hat\Sigma(x)^{-1} (X^{(1)})^\top\left(x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right)\epsilon^{(1)}.
\end{align*}
Next, we estimate each term using Lemma \ref{largedeviation} for random variables with bounded moment up to any order.
We first state several facts that will be commonly used in the proof.
By Fact \ref{fact_minv} (ii), we have that w.h.p. the operator norm of $X^{(1)}$ and $X^{(2)}$ are both bounded by $O(\sqrt{n})$.
Furthermore, the operator norm of $\hat\Sigma(x)^{-1}$ is bounded by $(x^2 + 1)^{-1} O(n_1 + n_2) = (x^2 + 1) O(p)$.
 % \be\label{op_X12}
%\|X^{(1)}\|\le \sqrt{(\sqrt{n_1} + \sqrt{p})^2 + n_1 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p}, \quad \|X^{(2)}\|\le \sqrt{(\sqrt{n_2} + \sqrt{p})^2 + n_2 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p},
% \ee
% and
% \be\label{op_Sig1}
%\| \hat\Sigma(x)^{-1}\|\le \frac1{x^2[(\sqrt{n_1} - \sqrt{p})^2 - n_1 \cdot p^{-c_{\varphi}}]+[(\sqrt{n_2} - \sqrt{p})^2 - n_2 \cdot p^{-c_{\varphi}}] }\lesssim \frac{1}{(x^2+1)p},
% \ee
%where we used that $3\le n_1/p\le \tau^{-1}$ and $3\le n_2/p\le \tau^{-1}$ for a small constant $\tau>0$.

@@

For the second step, we first argue that $\hat{x}$ is $O(1)$. Hence we can apply Claim \ref{claim_largedev1} to bound the concentration error of $g(x)$.
Using Fact \ref{fact_minv} (ii), we can obtain that with high probability,
\begin{align}
p(x^2+1)^{-1}&\lesssim \left( x^2 \left[(\sqrt{n_2}-\sqrt{p})^{2}-n_2p^{-c_\varphi}\right]^{-1} +\left[(\sqrt{n_1}-\sqrt{p})^{2}-n_1p^{-c_\varphi}\right]^{-1} \right)^{-1} \nonumber\\
&\preceq (X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1}(X^{(2)})^\top X^{(2)} =\left( x^2 [(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1} \right)^{-1} \nonumber\\
&\preceq \left( x^2 \left[(\sqrt{n_2}+\sqrt{p})^{2}+n_2p^{-c_\varphi}\right]^{-1} +\left[(\sqrt{n_1}+\sqrt{p})^{2}+n_1p^{-c_\varphi}\right]^{-1} \right)^{-1}\lesssim p(x^2+1)^{-1}, \label{op_uplow}
\end{align}
where we also used that $3\le n_1/p \le \tau^{-1}$ and $3\le  n_2/p \le \tau^{-1}$ for a small constant $\tau>0$. Then we get that
\be\label{eq_h1}h(1)= 2d^2 \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \hat\Sigma(1)^{-1}(X^{(2)})^\top X^{(2)}\right]\lesssim pd^2 .\ee
On the other hand, if $|x-1|\ge \delta$ for a small constant $\delta>0$, then we have
\begin{align}\label{eq_hA}
h(x)\gtrsim (1+x^2)\kappa^2 \cdot p\left( x^2 +1\right)^{-1} =p\kappa^2 ,
\end{align}
where in the first step we used that $|x-1|^2\gtrsim 1+x^2$ for $|x-1|\ge \delta$ and equation \eqref{op_uplow}.
and in the second step we used equation \eqref{choiceofpara0} \HZ{I'm not sure which condition we use? put it here}.
\begin{align}
g(x)&\ge  h(x) +\sigma^2(n_1+n_2-p) -p^{-1/2+c}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right) \nonumber\\
&\ge h(1) +\sigma^2(n_1+n_2-p) + p^{-1/2+c}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right) \ge g(1), \label{gA>g1}
\end{align}
under conditions $\sigma^2 =\OO( \kappa^2)$ and $d^2 =\oo({\kappa^2})$. %\eqref{choiceofpara0} \HZ{again, I don't see which condition we use here?}.
This gives that
Hence it remains to consider the region $|A-1|\le \delta$ for a small enough constant $\delta>0$.

First, using equations \eqref{op_uplow} and  \eqref{h'(A)1}, we obtain that for $1+3d^2/(2\kappa^2) \le x\le 1+\delta$,
$$h'(x)\ge \frac{2(x-1)\kappa^2 - 2d^2}{x^3} \cdot p^{-1} \tr\left[\left([(X^{(2)})^\top X^{(2)}]^{-1} + x^{-2} [(X^{(1)})^\top X^{(1)}]^{-1}\right)^{-1}\right] \gtrsim p\left(d^2 + |x-1|\kappa^2\right).$$
Thus we get that if $1+2d^2/\kappa^2 +p^{-1/4+c}\le x\le 1+\delta$,
\begin{align*}
h(x)-h(1)\ge h(x)-h\left(1+\frac{3d^2}{2\kappa^2}\right) \ge \int_{1+\frac{3d^2}{2\kappa^2}}^x h'(x)\dd x \gtrsim p\kappa^2 \cdot |x-1|^2  \gtrsim p\frac{d^4}{\kappa^2} + p^{1/2+2c} \kappa^2 .
\end{align*}
On the other hand, by Claim \ref{claim_largedev1} we have that
$$ g(x)=h(x)+\sigma^2 (n_1+n_2-p)+\OO \left(p^{1/2+c}\left(\sigma^2 +\kappa^2+d^2 \right)\right)$$
with high probability. Combining the above two estimates, we obtain that under the conditions $\sigma^2 =\OO( \kappa^2)$ and $d^2 =\oo({\kappa^2})$,
$$g(x)-g(1) = h(x)-h(1) + \OO \left(p^{1/2+c}\left(\sigma^2 +\kappa^2+d^2 \right)\right) >0$$
with high probability, i.e. $x$ cannot be the minimizer of $g$.
%Then if we choose the constant $c$ in Claim \ref{claim_largedev1} such that $e_1<2e_2$, we have
%$$ h(x)-h(1)\gtrsim  p^{-1/2+2e_2}\cdot p\kappa^2 \gg p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right),$$ where we use the assumption that $\sigma^2$ and $d^2$ are both at most $\OO(\kappa^2)$, which implies the inequality \eqref{gA>g1}.
Second, using equations \eqref{op_uplow} and  \eqref{h'(A)2}, we obtain that for $1-\delta \le x \le 1-3d^2/(2\kappa^2)$,
$$- h'(x)\ge  [2(1-x)\kappa^2 - 2xd^2] \cdot p^{-1} \tr\left[\left(x^2[(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1}\right)^{-1}\right]  \gtrsim p\left(d^2 + |x-1|\kappa^2\right).$$
Then with a similar argument as above, we can get that if $1-2d^2/\kappa^2 - p^{-1/4+c}\le x\le 1-\delta$, then
$$ h(x)-h(1) \gtrsim p\frac{d^4}{\kappa^2} +p^{1/2+2c} \kappa^2, $$
%under equation \eqref{choiceofpara0} \HZ{I don't see which condition we want to use here?},
which again implies that $g(x)>g(1)$, i.e. $x$ cannot be the minimizer of $g$. In sum, we obtain that the minimizer $\hat x$ of $g$ must satisfy equation \eqref{hatw_add1}.
%Next we prove the following estimate on the optimizer $\hat x$: with high probability,
%%\begin{lemma}
%%For the isotropic model, we have
%\be\label{hatv_add1}
%|\hat v - 1|= \OO\left(\cal E\right), \quad \cal E:=\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2} + p^{-1/2 + \e_0 /2+ 2\e}.
%\ee
%%\end{lemma}
%%\begin{proof}
% In fact, from the proof of Claim \ref{lem_hat_v} above, one can check that if $C\cal E \le |w-1| \le 2C\cal E$ for a large enough constant $C>1$, then $|h'(w)|\gtrsim \sigma^2$. Moreover, under equation \eqref{choiceofpara} we have
%$$h(w) =\OO(\sigma^2),\quad \text{for}\quad   |w-1|\le 2C\cal E.$$
%Thus we obtain that for $|w-1|\ge 2C\cal E$,
%$$\left|h(w) - h(\hat w)\right|\ge |h(w)-\min\{h(w_1),h(w_2)\}|\gtrsim \sigma^2 \cal E \gtrsim \cal E \cdot h(\hat w),$$
%which leads to $g(w) > g(\hat w)$ with high probability by equation \eqref{boundv-w}. Thus $w$ cannot be a minimizer of $g(v)$, and we must have $|\hat v - 1|\le 2C\cal E$. %Together with equation \eqref{hatw_add1}, we conclude equation \eqref{hatv_add1}.
%%\end{proof}
%
%{\cor require $\kappa^2\gg d^2 + p^{-1/2+c}\sigma^2$}
\end{proof}

\paragraph{Part $2$: a reduction to the bias and variance limits.}
Recall that the hard parameter sharing estimator $\hat{\beta}_2^{\MTL}$ is equal to $\hat{B} \hat{A}_2$.
The predication loss of hard parameter sharing is as follows
\begin{align}
L(\hat{\beta}_2^{\MTL}) &=\left\|(\Sigma^{(2)})^{1/2} \left( \hat B \hat A_2 - \beta^{(2)}\right)\right\| \nonumber\\
&=  \left\| (\Sigma^{(2)})^{1/2}\hat \Sigma(\hat x)^{-1} \left[(X^{(1)})^\top X^{(1)} (\hat x\beta^{(1)}-\hat x^2\beta^{(2)})+ (X^{(2)})^\top \epsilon^{(2)} + \hat x   (X^{(1)})^\top \epsilon^{(1)} \right]\right\|^2.\label{Lbeta_HPS}
\end{align}
Using Lemma \ref{lem_hat_v} and the concentration estimates in Lemma \ref{largedeviation}, we can simplify $L(\hat{\beta}_2^{\MTL})$ as follows.

\begin{claim}\label{claim_reduction}
Recall that $\hat \Sigma(1)$ is equal to $\hat \Sigma$ (cf. equation \eqref{def hatsig}).
In the setting of Proposition \ref{cor_MTL_loss}, we have the following estimate with high probability for any small constant $c>0$:
\begin{align*}
&\left|L(\hat{\beta}_2^{\MTL}) - \frac{2d^2}{p}\tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right] -\sigma^2  \tr\left({\hat \Sigma^{-1}  }\right)\right| \nonumber\\
\lesssim&  \frac{d^4 + \sigma^2 d^2}{\kappa^2}+p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2).
\end{align*}
\end{claim}
\begin{proof}
Our proof is divided into two steps. First, with the concentration estimates in Lemma \ref{largedeviation}, we can simplify $L(\hat{\beta}_2^{\MTL})$ as follows. With high probability, we have that for any small constant $c>0$,
\be\label{claim_largedev2} \left| L(\hat{\beta}_2^{\MTL})- \cal L(\hat x)\right| \le p^{-1/2+c} \left(\sigma^2 +\kappa^2 +d^2 \right),
\ee
where the function $\cal L(\hat x)$ is defined as
\begin{align*}
	\cal L(\hat x)	:=  &\hat x^2\left[(\hat x-1)^2 \kappa^2 +  (\hat x^2+1)d^2 \right] \cdot p^{-1}\bigtr{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma(\hat x)^{-1} \Sigma^{(2)} \hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} } \\
	& +\sigma^2\cdot \tr\left(\Sigma^{(2)}\hat \Sigma(\hat x)^{-1}\right).
\end{align*}
% \end{claim}
%We are now ready to finish the proof of Claim \ref{claim_reduction}.
We defer the proof of equation \eqref{claim_largedev2} until the end.
Next, we can further simplify $\cal L(\hat x)$ using Claim \ref{lem_hat_v} and $\Sigma^{(1)}=\Sigma^{(2)}=\id_{p\times p}$.
More precisely, we claim that with high probability
\begin{align}
&\left|\cal L(\hat x)- \frac{2d^2}{p} \tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right] -\sigma^2  \tr\left(\hat \Sigma^{-1}  \right)\right| \nonumber\\
\lesssim & \frac{d^4 + \sigma^2 d^2}{\kappa^2}+p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2). \label{simple12}
\end{align}
Combining equations \eqref{claim_largedev2} and \eqref{simple12}, we conclude Claim \ref{claim_reduction}.

>>>>

\paragraph{Part $1$: characterizing the global minimum of $f(A, B)$.}
Let $\hat{x}$ denote the global minimizer of $g(x)$.
We show that in the setting of Corollary \ref{cor_MTL_loss}, $\hat x$ is close to 1.
This gives us the global minimum of $f(A, B)$, since $\hat{B}$ is given by $\hat{x}$ using local optimality conditions.
First, we show that $g(x)$ and its expectation are close using standard concentration bounds.
\begin{claim}\label{claim_largedev1}
 In the setting of Corollary \ref{cor_MTL_loss}, for any $x$, we have that with high probability
$$\left| g(x) - \ex{g(x) \mid X^{(1)}, X^{(2)} }\right| \le p^{1/2+c} \left(\sigma^2 + \kappa^2+d^2 \right).$$
\end{claim}
\begin{proof}
 There are two terms in $g(A)$ from equation \eqref{eq_mtl_A12}.
 We will focus on dealing with the concentration error of the first term. The second term is similar to the first and we omit the details.
 For the first term, we expand into several equations under various situations involving the random noise and the random-effect model.
 \begin{align}
 & \left\| X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (x\beta^{(2)}-\beta^{(1)})+ \left(x^2 X^{(1)}\hat \Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right)\epsilon^{(1)} \right. \nonumber\\
		& \left. + x X^{(1)}\hat \Sigma(x)^{-1} (X^{(2)})^\top \epsilon^{(2)} \right\|^2=h_1(x) + h_2(x) +h_3(x) + 2h_4(x) + 2h_5(x) + 2h_6(x), \label{expand_6}
 \end{align}
 where
 \begin{align*}
&h_1(x):=  (\beta^{(1)}-x\beta^{(2)})^\top (X^{(2)})^\top X^{(2)}  \hat\Sigma(x)^{-1}  (X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (\beta^{(1)}-x\beta^{(2)})  ,\\
& h_2(x) := {(\epsilon^{(1)})^\top \left(x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right)^2\epsilon^{(1)}}  ,\\
& h_3(x):=  x^2 { (\epsilon^{(2)} )^\top X^{(2)} \hat\Sigma(x)^{-1}  (X^{(1)})^\top X^{(1)}\hat\Sigma(x)^{-1} (X^{(2)})^\top \epsilon^{(2)}} , \\
& h_4(x):=  (\epsilon^{(1)})^\top\left(x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right) X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (x\beta^{(2)}-\beta^{(1)}), \\
&h_5(x):=x (\epsilon^{(2)})^\top X^{(2)}\hat\Sigma(x)^{-1} (X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1} (X^{(2)})^\top X^{(2)} (x\beta^{(2)}-\beta^{(1)}),\\
&h_6(x):=x (\epsilon^{(2)})^\top X^{(2)}\hat\Sigma(x)^{-1} (X^{(1)})^\top\left(x^2 X^{(1)}\hat\Sigma(x)^{-1} (X^{(1)})^\top-\id_{n_1\times n_1}\right)\epsilon^{(1)}.
\end{align*}
Next, we estimate each term using Lemma \ref{largedeviation} for random variables with bounded moment up to any order.
We first state several facts that will be commonly used in the proof.
By Fact \ref{fact_minv} (ii), we have that w.h.p. the operator norm of $X^{(1)}$ and $X^{(2)}$ are both bounded by $O(\sqrt{n})$.
Furthermore, the operator norm of $\hat\Sigma(x)^{-1}$ is bounded by $(x^2 + 1)^{-1} O(n_1 + n_2) = (x^2 + 1)^{-1} O(p)$.
 % \be\label{op_X12}
%\|X^{(1)}\|\le \sqrt{(\sqrt{n_1} + \sqrt{p})^2 + n_1 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p}, \quad \|X^{(2)}\|\le \sqrt{(\sqrt{n_2} + \sqrt{p})^2 + n_2 \cdot p^{-c_{\varphi}}}\lesssim \sqrt{p},
% \ee
% and
% \be\label{op_Sig1}
%\| \hat\Sigma(x)^{-1}\|\le \frac1{x^2[(\sqrt{n_1} - \sqrt{p})^2 - n_1 \cdot p^{-c_{\varphi}}]+[(\sqrt{n_2} - \sqrt{p})^2 - n_2 \cdot p^{-c_{\varphi}}] }\lesssim \frac{1}{(x^2+1)p},
% \ee
%where we used that $3\le n_1/p\le \tau^{-1}$ and $3\le n_2/p\le \tau^{-1}$ for a small constant $\tau>0$.

@@

For the second step, using Claim \ref{claim_largedev1}, we have that $g(x)$ and $h(x)$ differ by at most $p^{1/2 + \e}(\sigma^2 + \kappa^2 + d^2)$.
Therefore, our goal reduces to showing that if $\hat{x}$ deviates too far from $1 \pm d^2 / \kappa^2$, it is no longer a global minimum of $g(x)$.
%we first argue that $\hat{x}$ is $O(1)$. Hence we can apply Claim \ref{claim_largedev1} to bound the concentration error of $g(x)$.
%Using Fact \ref{fact_minv} (ii), we can obtain that with high probability,
%\begin{align}
%p(x^2+1)^{-1}&\lesssim \left( x^2 \left[(\sqrt{n_2}-\sqrt{p})^{2}-n_2p^{-c_\varphi}\right]^{-1} +\left[(\sqrt{n_1}-\sqrt{p})^{2}-n_1p^{-c_\varphi}\right]^{-1} \right)^{-1} \nonumber\\
%&\preceq (X^{(1)})^\top X^{(1)} \hat\Sigma(x)^{-1}(X^{(2)})^\top X^{(2)} =\left( x^2 [(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1} \right)^{-1} \nonumber\\
%&\preceq \left( x^2 \left[(\sqrt{n_2}+\sqrt{p})^{2}+n_2p^{-c_\varphi}\right]^{-1} +\left[(\sqrt{n_1}+\sqrt{p})^{2}+n_1p^{-c_\varphi}\right]^{-1} \right)^{-1}\lesssim p(x^2+1)^{-1}, \label{op_uplow}
%\end{align}
%where we also used that $3\le n_1/p \le \tau^{-1}$ and $3\le  n_2/p \le \tau^{-1}$ for a small constant $\tau>0$. Then we get that
%\be\label{eq_h1}h(1)= 2d^2 \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \hat\Sigma(1)^{-1}(X^{(2)})^\top X^{(2)}\right]\lesssim pd^2 .\ee
%On the other hand, if $|x-1|\ge \delta$ for a small constant $\delta>0$, then we have
%\begin{align}\label{eq_hA}
%h(x)\gtrsim (1+x^2)\kappa^2 \cdot p\left( x^2 +1\right)^{-1} =p\kappa^2 ,
%\end{align}
%where in the first step we used that $|x-1|^2\gtrsim 1+x^2$ for $|x-1|\ge \delta$ and equation \eqref{op_uplow}.
%and in the second step we used equation \eqref{choiceofpara0} \HZ{I'm not sure which condition we use? put it here}.
%\begin{align}
%g(x)&\ge  h(x) +\sigma^2(n_1+n_2-p) -p^{-1/2+c}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right) \nonumber\\
%&\ge h(1) +\sigma^2(n_1+n_2-p) + p^{-1/2+c}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right) \ge g(1), \label{gA>g1}
%\end{align}
%under conditions $\sigma^2 =\OO( \kappa^2)$ and $d^2 =\oo({\kappa^2})$. %\eqref{choiceofpara0} \HZ{again, I don't see which condition we use here?}.
%This gives that
%Hence it remains to consider the region $|A-1|\le \delta$ for a small enough constant $\delta>0$.
We prove by contradiction.
First, suppose that $\hat{x} \ge 1 + 2d^2 /\kappa^2 + p^{-1/2 + \e}$.
For any $x \ge 1+3d^2/(2\kappa^2)$, we can lower bound the derivative of $h(x)$ using equation \eqref{h'(A)1} as follows
	\[ h'(x) \ge \frac{2(x-1)\kappa^2 - 2d^2}{x^3} \cdot p^{-1} \tr\left[\left([(X^{(2)})^\top X^{(2)}]^{-1} + x^{-2} [(X^{(1)})^\top X^{(1)}]^{-1}\right)^{-1}\right]
	\gtrsim p\kappa^2\cdot \frac{x-1}{x (1 + x^2)}. \]
Therefore, the difference between $h(x)$ and $h(1)$ is at least the following
\begin{align*}
		h(x)-h(1) \ge h(x) - h\left(1+\frac{3d^2}{2\kappa^2}\right)
	\ge \int_{1+\frac{3d^2}{2\kappa^2}}^x h'(x)\dd x \gtrsim p\kappa^2 \cdot \int_{1 + \frac{3d^2}{2\kappa^2}}^x \frac{x-1}{x(1 + x^2)} \dd x.
\end{align*}
When $\hat{x}$ is sufficiently far from $1$ (e.g. $2d^2 / \kappa^2 + p^{-1/4+\e}$), one can verify that $h(x) - h(1)$ is at least $\OO(p\frac{d^4}{\kappa^2} + p^{1/2+2c} \kappa^2) > \OO(p^{1/2 + c} (\sigma^2 + \kappa^2 + d^2))$, under the condition that $\sigma^2 =\OO( \kappa^2)$ and $d^2 =\oo({\kappa^2})$.
On the other hand, by triangle inequality and Claim \ref{claim_largedev1} we have that
\begin{align*}
	h(\hat{x}) - h(1) = g(\hat{x}) - g(1) + (h(\hat{x}) - g(\hat{x})) + (g(1) - h(1))
	\le \OO\left(p^{1/2+c}\left(\sigma^2 +\kappa^2+d^2 \right)\right).
\end{align*}
Hence, we have arrived at a contradition.

%Then if we choose the constant $c$ in Claim \ref{claim_largedev1} such that $e_1<2e_2$, we have
%$$ h(x)-h(1)\gtrsim  p^{-1/2+2e_2}\cdot p\kappa^2 \gg p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right),$$ where we use the assumption that $\sigma^2$ and $d^2$ are both at most $\OO(\kappa^2)$, which implies the inequality \eqref{gA>g1}.
Second, suppose that $\hat{x} \le 1 - 2d^2/ \kappa^2 - p^{-1/2 + \e}$.
Using equation \eqref{h'(A)2}, we obtain that for any $x \le 1 - 3d^2/(2\kappa^2)$,
	\[ -h'(x)\ge  [2(1-x)\kappa^2 - 2xd^2] \cdot p^{-1} \tr\left[\left(x^2[(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1}\right)^{-1}\right]  \gtrsim p\kappa^2 \cdot \frac{(1 - x)x^2}{1 + x^2}. \]
Using a similar argument to the first case, we get that the difference between $h(x)$ and $h(1)$ is at least the integral of the abvoe derivative.
%under equation \eqref{choiceofpara0} \HZ{I don't see which condition we want to use here?},
This implies that $\hat{x}$ cannot be too far from one.
Hence we have completed the proof.
%Next we prove the following estimate on the optimizer $\hat x$: with high probability,
%%\begin{lemma}
%%For the isotropic model, we have
%\be\label{hatv_add1}
%|\hat v - 1|= \OO\left(\cal E\right), \quad \cal E:=\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2} + p^{-1/2 + \e_0 /2+ 2\e}.
%\ee
%%\end{lemma}
%%\begin{proof}
% In fact, from the proof of Claim \ref{lem_hat_v} above, one can check that if $C\cal E \le |w-1| \le 2C\cal E$ for a large enough constant $C>1$, then $|h'(w)|\gtrsim \sigma^2$. Moreover, under equation \eqref{choiceofpara} we have
%$$h(w) =\OO(\sigma^2),\quad \text{for}\quad   |w-1|\le 2C\cal E.$$
%Thus we obtain that for $|w-1|\ge 2C\cal E$,
%$$\left|h(w) - h(\hat w)\right|\ge |h(w)-\min\{h(w_1),h(w_2)\}|\gtrsim \sigma^2 \cal E \gtrsim \cal E \cdot h(\hat w),$$
%which leads to $g(w) > g(\hat w)$ with high probability by equation \eqref{boundv-w}. Thus $w$ cannot be a minimizer of $g(v)$, and we must have $|\hat v - 1|\le 2C\cal E$. %Together with equation \eqref{hatw_add1}, we conclude equation \eqref{hatv_add1}.
%%\end{proof}
%
%{\cor require $\kappa^2\gg d^2 + p^{-1/2+c}\sigma^2$}
\end{proof}

\paragraph{Part $2$: a reduction to the bias and variance limits.}
Recall that the hard parameter sharing estimator $\hat{\beta}_2^{\MTL}$ is equal to $\hat{B} \hat{A}_2$.
Using the local optimility condition for $\hat{B}$, we obtain the predication loss of HPS as follows
\begin{align}
L(\hat{\beta}_2^{\MTL}) &=\left\|(\Sigma^{(2)})^{1/2} \left( \hat B \hat A_2 - \beta^{(2)}\right)\right\| \nonumber\\
&=  \left\| (\Sigma^{(2)})^{1/2}\hat \Sigma(\hat x)^{-1} \left[(X^{(1)})^\top X^{(1)} (\hat x\beta^{(1)}-\hat x^2\beta^{(2)})+ (X^{(2)})^\top \epsilon^{(2)} + \hat x   (X^{(1)})^\top \epsilon^{(1)} \right]\right\|^2.\label{Lbeta_HPS}
\end{align}
Using Lemma \ref{lem_hat_v} and the concentration estimates in Lemma \ref{largedeviation}, we simplify $L(\hat{\beta}_2^{\MTL})$ as follows.

\begin{claim}\label{claim_reduction}
Recall that $\hat \Sigma(1)$ is equal to $\hat \Sigma$ (cf. equation \eqref{def hatsig}).
In the setting of Claim \ref{cor_MTL_loss}, we have the following estimate w.h.p.
\begin{align*}
&\left|L(\hat{\beta}_2^{\MTL}) - \frac{2d^2}{p}\tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right] -\sigma^2  \bigtr{{\hat \Sigma^{-1}  }}\right| \nonumber\\
\lesssim&  \frac{d^4 + \sigma^2 d^2}{\kappa^2}+p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2).
\end{align*}
\end{claim}
\begin{proof}
Our proof is divided into two steps. First, using Lemma \ref{largedeviation}, we show that
\be\label{claim_largedev2} \left| L(\hat{\beta}_2^{\MTL})- \cal L(\hat x)\right| \le p^{-1/2+c} \left(\sigma^2 +\kappa^2 +d^2 \right),
\ee
where $\cal L(\hat x)$ is defined as
\begin{align*}
	\resizebox{0.97\hsize}{!}{%
	$\cal L(\hat x)	:=  \hat x^2\left[(\hat x-1)^2 \kappa^2 +  (\hat x^2+1)d^2 \right] \cdot p^{-1}\bigtr{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma(\hat x)^{-1} \Sigma^{(2)} \hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} }
	 +\sigma^2\cdot \bigtr{\Sigma^{(2)}\hat \Sigma(\hat x)^{-1}}.$
	}%
\end{align*}
% \end{claim}
%We are now ready to finish the proof of Claim \ref{claim_reduction}.
Next, we further simplify $\cal L(\hat x)$ since $\hat{x}$ is close to one and $\Sigma^{(1)},\Sigma^{(2)}$ are both isotropic
\begin{align}
	&\left|\cal L(\hat x)- \frac{2d^2}{p} \tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right] -\sigma^2  \bigtr{\hat \Sigma^{-1}}\right| \nonumber\\
\lesssim & \frac{d^4 + \sigma^2 d^2}{\kappa^2}+p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2). \label{simple12}
\end{align}
Combining equation \eqref{claim_largedev2} and \eqref{simple12}, we obtain the desired claim.
We prove these two equations one by one as follows.
